Leadership across cultures and contexts: an exploratory analysis

### 2020-05-14 leadership_across_cultures_contexts.Rmd
Tweaked upset plot stuff; edited text on human capital. Finished conclusion. Started revising abstract.

Leader-follower dynamics are central to much of social life. Processes of leadership and followership observed in nonindustrial societies and smaller-scale organizations are often distinct from those observed among WEIRD societies and their complex organizations. Many researchers have turned to evolutionary theory to integrate diverse perspectives on human leadership and to better understand the diversity of leadership and followership observed across human cultures and across social contexts. Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved. Currently, though, there is limited systematic data on the nature of leadership in such societies.

We build on our previous cross-cultural findings to systematically evaluate dinebnsions of leadership across the full range of human cultural diversity. Over 1200 ethnographic texts from the Human Relations Area Files (HRAF) were coded on leader qualities, their functions, and the costs and benefits for both leaders and followers across diverse contexts. We discovered evidence for both cultural universals in leadership, as well as important variation with this sample of largely "non-WEIRD" populations. Leaders in the ethnographic record are most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial. The primary benefits leaders offer to followers are resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources. The primary benefits leaders receive from followers include material resources, increased social status, and social services from followers. The most frequent costs of leadership for leaders include loss of social status, resources or social services and increased risk of harm or conflict. For followers, the most frequent costs include loss of resources, costs in providing social services, and increased risk of harm or conflict.

@@

## The five-fold way OR Leadership markets

In our view, the dominance-prestige and "dual-model" theories of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers". In short, these models have conflated "benefits" with knowledge and expertise, and "costs" with physical formidability. @van_vugt_dual_2019, for instance, state that the source of deference for prestige-style leadership is "Information asymmetry", and for dominance-style, "Power asymmetry" (their Table 1). Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].

We therefore tenatively advance an alternative scheme whereby individuals possess some combination of knowledge/expertise and physical formidability which they use to provide benefits and impose costs. If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016]. Hence, we should be able to find examples of leaders with various combinations of these five elements.

In our data, we used the high status quality as "High status", the prosociality feature as "Provide benefits", any one of punishment, aggressive, feared, and killer as "Impose costs", knowledge/intellect and/or experience/accomplished as "Knowledge", and physically formidible and/or physically healthy as "Physical". We then calculated the fraction of text records with evidence for one or more of the five elements. See Figure \@ref(fig:fivefold).

```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic dimensions of leadership among all text records (an 'upset' plot) [@gehlenborg2019]."}
upset(df_fivefold, order.by='freq', nintersects = 19, text.scale = 2)
```

As Figure \@ref(fig:fivefold) reveals,

<!-- DAVIDS MARKET FORCES IDEAS -->
We agree with @pietraszewski_evolution_2019 that leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019], however, much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination. We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms. 
 <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- CAN'T JUST BE NONWEIRD -->
In recognizing the limited generalizability of most extant psychological, sociological, and managerial data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale". At the same time, anthropologists have made the opposite mistake in often failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].

<!-- OUR SUPPORT FOR THEORIES -->
Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions. We found leaders in the ethnographic record are portrayed as generally prosocial, are equipped with special competencies, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources. These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].

<!-- SHAMANS AND BEYOND DUAL MODEL -->
Shamans emerged as an important category of leaders that are rarely discussed in the evolutionary literature [but see @singh_cultural_2017]. Supernatural qualities were among the top ten leader qualities across cultures (Figure \@ref(fig:varsupport)). Shamans share qualities with prestige-style leaders, such as having specialized knowledge and abilities, yet appear, at least at times, to use their knowledge to instill fear, similar to dominance-style leaders.

[@hawley_ontogenesis_1999; @hawley_myth_2008]

Leaders vs. Status vs. Provide benefits vs. Impose costs vs. Information vs. Physical

We therefore suggest moving beyond dichotomous, or "dual-process", models of leadership.

and focusing on inter-individual variation in broader typologies of leadership competencies. There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies. The degree to which a leader relies on such competencies will be shaped by their individual levels of them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms. For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies. Cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016] and the ecology of leadership dimensions can be better captured than dichotomous frameworks offer.

Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.

<!-- DIVERSE FUNCTIONS -->
Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014]. Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group. However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].

<!-- LEADER SUCCESS AND INEQUALITY -- REPETITIVE ? -->
The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.

<!-- CONNECT TO WERID -->
Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature. Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriented* framework of managerial behaviors (i.e., our Management vs. Prosociality features of leader functions), however, our results do not support the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic] (although charisma was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%). See Figure \@ref(fig:cluster).

<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization. Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers. Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups. Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history. Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018]. Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].

<!-- CLOSING STATEMENT -->
Leaders likely rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms. Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics.

@@

>>>>

Processes of leadership and followership, which are central to social life in all societies, nevertheless vary from the informal roles observed in nonindustrial societies and smaller-scale organizations to the formal roles observed among WEIRD societies and their large-scale, complex organizations. Many researchers have turned to evolutionary theory to better understand this diversity. Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved. Currently, though, there is limited systematic data on the nature of leadership in such societies.

We build on our previous cross-cultural findings to systematically evaluate dimensions of leadership across a broad range of human cultural diversity. Over 1200 ethnographic texts from 59 mostly non-WEIRD populations in Human Relations Area Files (HRAF) were coded on leader qualities and functions. Because adaptations for leadership and followership could only have evolved if the fitness benefits exceeded the costs, we coded the costs and benefits for both leaders and followers across diverse contexts. We discovered evidence for both cultural universals in leadership, as well as important variation. Leaders in the ethnographic record were most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial. The primary benefits leaders offered to followers were resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources. The primary benefits leaders received from followers included material resources, increased social status, and social services. The most frequent costs of leadership included loss of social status, resources or social services and increased risk of harm or conflict. For followers, the most frequent costs included loss of resources, costs in providing social services, and increased risk of harm or conflict.

Shamans emerged as an important category of leaders who do not clearly conform to influential models that emphasize two leader strategies: using knowledge and expertise to provide benefits to followers vs. using physical formidability to impose costs. Instead, shamans and other leaders with supernatural abilities use their knowledge to both provide benefits and impose costs on others. We therefore propose a modified scheme in which leaders deploy their cognitive, social, and somatic capital to provide benefits and/or impose costs on others.

@@

## Leadership and cognitive, social, and somatic capital

In our view, the dominance-prestige and "dual-model" theories of leadership, while advancing the discipline, cannot easily account for shaman leaders nor "managers". These models, in short, have conflated the ability to provide "benefits" with knowledge and expertise, and the ability to impose "costs" with physical formidability. @van_vugt_dual_2019, for instance, state that the source of deference for prestige-style leadership is "Information asymmetry", and for dominance-style, "Power asymmetry" (their Table 1). Instead, as shaman leaders demonstrate, knowledge and expertise can both provide benefits and impose costs, and the same goes for physical formidability. The use of knowledge to impose costs is widely recognized by many social scientists. The role of gossip and other forms of indirect aggression in social dominance and achieving leadership roles, for instance, has been extensively investigated in WEIRD populations [e.g., @hawley_ontogenesis_1999; @hess_gossip_2017; @hess_sex_2006].

We therefore tenatively advance an alternative scheme, similar to @von_rueden_roots_2014[] [and inspired by @kaplan_embodied_2003; @kaplan_evolutionary_2009; @borgerhoff_mulder_intergenerational_2009], whereby individuals possess some combination of cognitive capital, social capital, and somatic capital that they use to provide benefits and impose costs. If their abilities to provide benefits or impose costs are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019]. Hence, we should be able to find examples of leaders with various combinations of these five elements.^[We omit material capital because this could be both a means to leadership as well as a benefit of leadership.]

To test this idea, we operationized the five elements using the leadership qualities cluster analysis (Figure \@ref(fig:cluster)A) as guidance (see Table \@ref(tab:leadermarket)). We then calculated the fraction of text records with evidence for one or more of the five elements. See Figure \@ref(fig:fivefold).

```{r fivefold, fig.width=10, fig.cap="The intersections of evidence for our five basic dimensions of leadership among all text records (an 'upset' plot) [@gehlenborg2019]."}
upset(df_fivefold2, order.by = 'freq', point.size = 5, text.scale = 2, nintersects = 18)
```

As Figure \@ref(fig:fivefold) reveals, descriptions of leadership in the ethnographic record includes a heterogeneous mix of these traits. In particular, not all leaders have evidence for high *Social capital*, leaders with *Cognitive capital* provide benefits *and* impose costs, and *Somatic capital* is rarely mentioned, but when it is, it is associated with providing benefits as well as imposing costs.

<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019]. Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination. We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.  -->
 <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- CAN'T JUST BE NONWEIRD -->
In recognizing the limited generalizability of most extant psychological, sociological, and managerial data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale". At the same time, anthropologists have made the opposite mistake in often essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].

<!-- OUR SUPPORT FOR THEORIES -->
Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs. We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources. These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].

<!-- SHAMANS AND BEYOND DUAL MODEL -->
Shamans emerged as an important category of leaders that are rarely discussed in the evolutionary literature [but see @singh_cultural_2017]. Supernatural qualities were among the top ten leader qualities across cultures (Figure \@ref(fig:varsupport)). Shamans share qualities with prestige-style leaders, such as having specialized knowledge and abilities, yet appear, at least at times, to use their knowledge to instill fear, similar to dominance-style leaders. We therefore suggest moving beyond dichotomous or "dual" models of leadership, which have tended to conflate knowledge and expertise with providing benefits. Instead, our data indicate that leaders employ cognitive, social, and somatic capital to provide benefits and impose costs (Figure \@ref(fig:fivefold)). This view aligns with the theoretical framework developed by many scholars that emphasizes the importance of embodied capital (neural and somatic) and social and material capital in the evolution of human social organization [@kaplan2003neural; @kaplan_embodied_2003; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @bowles_emergence_2010].

<!-- and focusing on inter-individual variation in broader typologies of leadership competencies. There is likely to be substantial overlap in, and interaction between physical, supernatural, cognitive, and other prosocial competencies. The degree to which a leader relies on such competencies will be shaped by their individual levels of them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms. For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies. Cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016] and the ecology of leadership dimensions can be better captured than dichotomous frameworks offer. -->

<!-- DIVERSE FUNCTIONS -->
<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014]. Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group. However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019]. -->

<!-- Furthermore, higher-order explanatory variables such as "subsistence strategy" are probably less relevant for understanding the complexity of behavioral diversity than more salient group contexts.  -->

<!-- LEADER SUCCESS AND INEQUALITY -- REPETITIVE ? -->
<!-- The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership. -->

<!-- CONNECT TO WERID -->
Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Wofford's *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature. Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriented* framework of managerial behaviors (i.e., our Management vs. Prosociality features of leader functions). However, our results, which found evidence for oratory skills in about 50% of cultures and charisma in about 20%, provide only partial support for the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic]. See Figure \@ref(fig:cluster).

<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
<!-- These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization. Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers. Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups. Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history. Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].  -->

<!-- CLOSING STATEMENT -->
Leaders rely a range of individual capacities, including physical, supernatural, and cognitive endowments, as well as their social networks, in order to organize group members, implement strategic actions, and provide prosocial services to the group, all while conforming to cultural norms. Dichotomizing this complexity into two distinct social strategies, although theoretically and empirically convenient, likely fails to capture much of the ethnographic reality of human leadership dynamics. Given the broad range of leadership dimensions represented in the ethnographic record, we suggest leaders-follower dynamics should be increasingly incorporated in evolutionary models behavior and group dynamics, including demography, social complexity and inequality, cooperation and punishment, and resource exchange [c.f., @bird_variability_2019; @ember_our_2018; @micheletti_modelling_2020; @smith_cultural_2020; @birch2017philosophy; @marlowe_more_2008; @henrich_markets_2010; @mace_co-evolution_1998; @mace_phylogenetic_2005; @mattison_evolution_2016].

@@

## Leader market variable operationalization

```{r leadermarket}

frmt <- function(label, vars){
  c(label, rep('', length(vars) - 1))
}

marketvars <-
  tibble(
    Element = c(
      'High status', 
      frmt('Cognitive capital', cognitive_vars),
      frmt('Somatic capital', physical_vars),
      frmt('Provide benefits', benefit_ability_vars),
      frmt('Impose costs', cost_ability_vars)
    ),
    `Component variables` = c(
    'High status',
    names(cognitive_vars),
    names(physical_vars),
    names(benefit_ability_vars),
    names(cost_ability_vars)
    )
  )
kable(marketvars, caption = 'Operationalization of our leadership components')
```
